---
title: "Multiscale entropy"
output:
  pdf_document: default
  html_notebook: default
---
Written by Dr. John A.E. Anderson in collaboration with Dr. John G. Grundy
09/12/2020

This is an attempt to write a multiscale entropy notebook based on the MSEntropy package in R (https://github.com/jcaude/MSEntropy)

The first step is to read in the files matching a specific criteria (condition) using the readbulk package - we can then compute entopy within each participant. 

First load the packages necessary for the project...
```{r,comment=FALSE,message=FALSE,error=FALSE}
library(MSEntropy)
library(dplyr)
library(reshape2)
library(readbulk)
library(readr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(scales)
library(eegUtils)
library(colorspace)
library(patchwork)
```

Now read in your data...
try keeping only the first two electrodes to make sure that grouping is electrode *within* subject not vice-versa
```{r,comment=FALSE,message=FALSE,error=FALSE,echo=FALSE, warning=FALSE}
#first read all the files into a dataframe using the "readbulk package"
#the subdirectories =TRUE function means that we expect every person's data to be in a separate subdirectory. We're using the read_table2 function since our data is tab-separated, but we could easily adapt this to read CSV files etc.

temp <- read_bulk(directory = "/Users/john/Downloads/MSE_shedden_2019/BE_ERP_text_files/test2/", extension =".txt", subdirectories = TRUE,fun = read_table2)
#remove columns that aren't electrodes
temp$X69 <- NULL
#reshape the dataset from wide to long format so that we can calculate entropy by variable (electrode)
temp2 <- melt(temp, id.vars=c("File", "Subdirectory"))
#convert the value from character or whatever to numeric
temp2$value <- as.numeric(temp2$value)

#now we want to grab the file names which uniquely identify each person and condition
files <- unique(temp$File)
electrodes <- colnames(temp)
electrodes <- head(electrodes,-2)
#now get a list of files and electrodes
files_electrodes <- expand.grid(electrodes, files)
files_electrodes_combo <-paste(files_electrodes$Var1, files_electrodes$Var2,sep = "_") 
```
Now calculate MSE by condition/participant
```{r,comment=FALSE,message=FALSE,error=FALSE,echo=FALSE, warning=FALSE}
#this function calculates MSE using scales 1:20 with an r of 0.5 - change these values if you are interested in looking at this. We've specified that input$subdirectory is NULL and input$File is NULL which removes these non-numeric inputs before calculating MSE

#sets number of scales (normally 20)
scalenum=20

MSE_func <- function(input,scalenum){
  input$Subdirectory <- NULL
  input$File <- NULL
  input$variable <- NULL
  MSEntropy::MSE(x = as.matrix(input),scales = 1:scalenum,r = 0.5,m = 2)
}

#Finally, we apply the MSE function we defined above to the whole data frame (temp), but specify that the function should be applied within each File (the group_by File command). We then specify that we want the output to be a dataframe and that the names should be the unique file identifiers.
MSE_out <- temp2 %>%
  group_by(File, variable) %>% #Make sure that the grouping is within electrode and within subject (variable = electrode)
  group_map(~ MSE_func(.x, scalenum)) %>%
  data.frame()%>% #everything up to here works...
  setNames(files_electrodes_combo) %>%
  rename_all(
    funs(stringr::str_replace_all(., ".txt", ""))
    )
#define the bins using the length of the data structure (sequence along from 1 to length of data frame & add this variable to the output)
MSE_out$bin <- seq(from=1, to=length(MSE_out[,1]))
#now melt the dataframe (go to long format) to make it easier to split the subject ID and variable information
MSE_out_long <- melt(MSE_out,id.vars = "bin")%>% 
  separate(variable, "_",
                into = c("Electrode", "ID", "MSE","test","date","context","Cond","accuracy"), 
                remove = FALSE)
#	MSE_out_long$ID = substr(MSE_out_long$ID_Cond,1,4)
#	MSE_out_long$Cond = substr(MSE_out_long$ID_Cond,5,9)
	
#now put the data into a format that can be analyzed (bin and condition by ID)	
MSE_out_long <- subset(MSE_out_long, select=c("ID","Cond", "context","Electrode","bin","value"))
(MSE_out_wide <- dcast(data = MSE_out_long, formula = ID + Cond + context ~ Electrode + bin, value.var = "value"))
	
#test plotting function
fz <- subset(MSE_out_long, Electrode == "AF8")
ggplot(fz, aes(x=bin, y = value, color=Cond, group=Cond))+
  stat_summary(
    geom = "point",
    fun.y = "mean",
    size = 3,
    shape = 24,aes(col=Cond)
  )	+ 
  scale_color_manual(values=c("darkred","darkgreen"))+facet_wrap( ~ context)+theme_pubr()
```
PLS analysis

```{r}
#
library(TExPosition)
library(TInPosition)
library(data4PCCAR)


X <- subset(MSE_out_wide, select=c("Cond","context"))
X <- fastDummies::dummy_cols(X)
X$Cond<- NULL
X$context <- NULL
Y <- MSE_out_wide
Y$ID <- NULL
Y$Cond<-NULL
Y$context<-NULL

pls.mse <- tepPLS(DATA1 = Y, DATA2 = X, center1 = TRUE, scale1 = TRUE, center2 = TRUE, scale2 = TRUE,graphs = FALSE, k = 2)


#now using the bootstrap ratios from the Boot4PLSC function in the data4PCCAR package
#Dataset 1 (fISO) is represented by "i" in the output, while dataset 2 (Tau) is represented by j in the output
pls.boot <- Boot4PLSC(X, Y, center1 = TRUE, center2 = TRUE,
  scale1 = "ss1", scale2 = "ss1", Fi = NULL, Fj = NULL,
  nf2keep = 3, nIter = 1000, critical.value = 2, eig = FALSE,
  alphaLevel = 0.05)

pls.perm <- perm4PLSC(X, Y, center1 = TRUE, center2 = TRUE,
  scale1 = "ss1", scale2 = "ss1", nIter = 1000, permType = "byMat",
  compact = FALSE)
pls.perm$pOmnibus


```

Plotting the results...

```{r default-geom, fig.height = 4, fig.width = 4}
#coordinates for the EEG plot
EEG_64_coords <- read_csv("EEG_64_coords.csv")

#for the function below - choose the LV number you want to plot.
brain_fig <- function(lv_numeric){
LV <-as.data.frame(pls.boot$bootRatios.j[,lv_numeric])
LV <- tibble::rownames_to_column(LV, "Electrode")%>% 
  separate(Electrode, "_",
                into = c("Electrode", "bin")) %>%
  `colnames<-`(c("Electrode","bin","BSR"))
#merge the Lv1 bootstrap ratio data with the column indicating reliability
Lv_rel <- as.data.frame(pls.boot$bootRatiosSignificant.j[,lv_numeric])
Lv_rel <- tibble::rownames_to_column(Lv_rel, "Electrode")%>% 
  separate(Electrode, "_",
                into = c("Electrode", "bin")) 
output <- merge(LV, EEG_64_coords, by ="Electrode")
output <- merge(output, Lv_rel, by =c("Electrode","bin"))%>%
  `colnames<-`(c("Electrode","bin","BSR","x","y","z","reliable"))
output$bin <- as.factor(as.numeric(output$bin)) 
return(assign(paste("Lv",lv_numeric, sep=""), output))
}

Lv1 <- brain_fig(1)
Lv2 <- brain_fig(2)


brain_plot_LV1 <- ggplot(Lv1,
       aes(x = x,
           y = y,
           fill = BSR,
           label = Electrode)) +
  geom_topo(grid_res = 200,
            interp_limit = "head",
            chan_size=0.5,head_size = 0.5)+ 
  #scale_fill_continuous_diverging(palette = "Blue-Red 3", breaks=c(-4,-2,0,2)) + 
    scale_fill_viridis_c(breaks=c(-4,-2,0,2))+
  theme_void() + 
  coord_equal() + 
  labs(fill = expression(paste("BSR"))) + 
  facet_wrap(~bin)+
  theme_void()+ 
  ggtitle('Plot 1')

brain_plot_LV2 <- ggplot(Lv2,
       aes(x = x,
           y = y,
           fill = BSR,
           label = Electrode)) +
  geom_topo(grid_res = 200,
            interp_limit = "head",
            chan_size=0.5,head_size = 0.5)+ 
  #scale_fill_continuous_diverging(palette = "Blue-Red 3", breaks=c(-4,-2,0,2)) + 
    scale_fill_viridis_c(breaks=c(-4,-2,0,2))+
  theme_void() + 
  coord_equal() + 
  labs(fill = expression(paste("BSR"))) + 
  facet_wrap(~bin)+
  theme_void()+ 
  ggtitle('Plot 3')

#doing this with geom tile
cols=c("gray","black")

Tile_Plot_LV1 <- ggplot(Lv1, aes(x = bin, y= Electrode)) +
    geom_tile(aes(fill = BSR, colour=factor(reliable), size=factor(reliable)))+ #alpha=reliable==FALSE
    #scale_fill_continuous_diverging(palette = "Blue-Red 3", breaks=c(-4,-2,0,2))+
  scale_fill_viridis_c(breaks=c(-4,-2,0,2))+
  #scale_alpha_manual(values = c(1, 0.5), guide = FALSE)+
    scale_colour_manual(values=cols)+
    scale_size_manual(values=c(0,0.5))+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(legend.position = "none") + 
  ggtitle('Plot 2')

Tile_Plot_LV2 <- ggplot(Lv2, aes(x = bin, y= Electrode)) +
    geom_tile(aes(fill = BSR, colour=factor(reliable), size=factor(reliable)))+ #alpha=reliable==FALSE
    #scale_fill_continuous_diverging(palette = "Blue-Red 3", breaks=c(-4,-2,0,2))+
  scale_fill_viridis_c(breaks=c(-4,-2,0,2))+
  #scale_alpha_manual(values = c(1, 0.5), guide = FALSE)+
    scale_colour_manual(values=cols)+
    scale_size_manual(values=c(0,0.5))+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(legend.position = "none") + 
  ggtitle('Plot 4')
  


ggarrange(ggarrange(brain_plot_LV1, brain_plot_LV2, nrow=2), ggarrange(Tile_Plot_LV1, Tile_Plot_LV2, nrow=2), ncol=2, widths = c(2,1))
#if screen freezes: dev.set(dev.next())

##### using the topoplot function
#topoplot(data=subset(Lv2, bin==5),
#         palette = "magma",
#         highlights = c("B5"),
#         interp_limit = "head",limits = c(-3,3))
```
